---
title: "Sta 522 HW7"
author: "Daniel Truver"
date: "3/28/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#### (1) NCVS

```{r setup1}
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
ncvs = read.csv("ncvs2000.csv")
```

##### (a) Random Assignment

```{r}
n = nrow(ncvs)
treatment = sample(c(rep(0, n/2), rep(1, n/2)))
ncvs$treatment = treatment
```

##### (b) Example Variables

```{r didItWork}
treat.df = ncvs %>%
  filter(treatment == 1)
control.df = ncvs %>% 
  filter(treatment == 0)
treat.avg = data.frame(c(
  mean(treat.df$age),
  mean(treat.df$married == 1),
  mean(treat.df$race == 1),
  mean(treat.df$race == 2),
  mean(treat.df$race == 3),
  mean(treat.df$race == 4),
  mean(treat.df$sex == 0),
  mean(treat.df$sex == 1)
))
rownames(treat.avg) = c("age", "married", "White", "Black", "Native American",
                        "Asian", "male", "female")
control.avg = data.frame(c(
  mean(control.df$age),
  mean(control.df$married == 1),
  mean(control.df$race == 1),
  mean(control.df$race == 2),
  mean(control.df$race == 3),
  mean(control.df$race == 4),
  mean(control.df$sex == 0),
  mean(control.df$sex == 1)
))
rownames(control.avg) = rownames(treat.avg)
```

```{r echo=FALSE}
df.avg = cbind(treat.avg, control.avg)
colnames(df.avg) = c("treatment", "control")
knitr::kable(df.avg, col.names = c("Treatment", "Control"),
             caption = "Sample averages for treatment and control groups")
```

##### (c) Discussion

Every variable is pretty much the exact same. Let's add a column of differences to the table above and take a look at the distribution of the one continuous variable, age.

```{r, echo=FALSE}
df.avg$diff = df.avg$treatment - df.avg$control
knitr::kable(df.avg, col.names = c("Treatment", "Control", "Difference"),
             caption = "Differences between treatment and control groups")
```

```{r, echo=FALSE}
{
  par(mfrow = c(1,2))
  hist(treat.df$age)
  hist(control.df$age)
}
```

#### (2) Completely Randomized, Two Sample Experiment

```{r echo=FALSE}
data.2 = data.frame(individual = 1:4,
                    Y.1 = c(5,2,NA,NA),
                    Y.0 = c(NA,NA,1,3))
```

##### (a) Questionable Entries

The question marks are present because we did not observe the values for individuals 3 and 4 under the treatment, and we did not observe the values for individuals 1 and 2 under the control. Within one run of this experiment, is impossible to obtain both treatment and control observations for an individual.

##### (b) Fisher Sharp Null

Under the Fisher Null hypothesis, the data would look like this.

```{r echo=FALSE}
Y = c(5,2,1,3)
n = length(Y)
data.2 %>% mutate(Y.1 = Y, Y.0 = Y) %>%
  knitr::kable(., col.names = c("individual", "Y_i(1)", "Y_i(0)"),
               caption = "Data under Fisher sharp null hypothesis")
```

##### (c) Possible Assignments

```{r}
trt = data.frame(combn(data.2$individual, m = 2))
```

```{r, echo=FALSE}
knitr::kable(trt, col.names = paste0("treatGroup", 1:6),
             caption = "Possible assignments to treatment group")
```  

```{r}
y1 = rep(NA, 6)
y0 = rep(NA, 6)
for (i in 1:6){
  treatGroup = data.2$individual %in% trt[,i]
  y1[i] = mean(Y[treatGroup])
  y0[i] = mean(Y[!treatGroup])
}
diff.avg = y1-y0
```

##### (d) Fisher p-value

```{r, echo=FALSE}
ggplot(data = data.frame(diff.avg), aes(x = diff.avg)) +
  geom_histogram() +
  geom_vline(xintercept = 3/2, col = "red")
```

The probability is 0.33 that we would see a value as or more extreme than the observed value by random chance.

##### (e) 

```{r}
treat = c(5,2)
contr = c(1,3)
full = c(treat, contr)
conditions = c(1,1,0,0)
S = 10000
y.bar = rep(NA, S)
for (i in 1:S){
  samp = sample(conditions)
  y.bar[i] = mean(full[samp == 1]) - mean(full[samp == 0])
}
```

```{r, echo=FALSE}
{
  hist(y.bar, xlab = "Treatment effects")
  abline(v=3/2, col = "red")
}
mean(y.bar >= 3/2) 
```

Our computed p-value here is `r mean(y.bar >= 3/2)`. It is nearly identical to the one computed in part (d).